---
title: "5513 - Assignment 3"
format: html
editor: visual
---

## Executive Summary

Social media addiction may quietly affect students' ability to manage their time and emotions. This report explores the link between addiction levels and self-control using 2024 survey data. Students with higher addiction scores often showed significantly lower self-control. Based on these findings, it is recommended to provide proactive support through school programs and awareness workshops.

## Introduction

Social media has become essential to students’ daily routines, often influencing how they spend their time and interact with others. While it provides entertainment and connection, growing dependence on these platforms can lead to addictive behaviours. One area of interest is whether frequent and potentially addictive social media use might influence students’ ability to manage other important aspects of life, such as academics, relationships, and personal responsibilities, which are often associated with self-control. However, there are few references on whether heavy social media use interferes with these behaviours or not. To explore this, we investigate whether higher levels of social media addiction are associated with lower levels of self-control among students. Using survey data collected in 2024, we analyse responses related to addiction levels and behavioural outcomes. The findings aim to reveal how digital habits impact student well-being. This may help inform strategies to promote healthier online behaviour.

## Methodology

### 1. Analytical Approach

Our research question is: Will the addiction to social media reduce student's sense of self-control?

To answer this question, we used the [Students Social Media Addiction](https://www.kaggle.com/datasets/adilshamim8/social-media-addiction-vs-relationships) data set. We selected 4 variables from it for analysis, We divided the Addicted_Level into 3 groups based on their level of social media addiction (Low, Medium, High), and analyzed the self-control ability through 3 aspects of each group.

### 2. Data description

The details of the dataset’s variables and their descriptions:

Student_ID: Unique identifier for each student

Addicted_Level: Level of social media addiction of the student (1-3: Low, 4-6: Medium, 7-10: High）

Avg_Daily_Usage_Hours： Number of hours students spends using social media per day

Conflicts_Over_Social_Media： Reported number of arguments or disagreements student have with their family, friends and/or partners

Affects_Academic_Performance： Student have had their academic result impacted by using social media (1: Yes, 0: No)

### 3. Cleaning Data {#sec-cleaning}

```{r}
library(dplyr)
library(readr)
raw_data <- read_csv("data/Students_Social_Media_Addiction.csv")
processed_data <- raw_data %>%
  mutate(Addicted_Level = case_when(
            Addicted_Score <= 3 ~ "Low",
            Addicted_Score <= 6 ~ "Medium",
            TRUE ~ "High"
        ),
        Addicted_Level = factor(Addicted_Level, levels = c("Low", "Medium", "High")),
    Affects_Academic_Performance = ifelse(Affects_Academic_Performance == "Yes", 1, 0)
        ) %>%
  select(Student_ID,
         Addicted_Level,
         Avg_Daily_Usage_Hours, 
         Conflicts_Over_Social_Media, 
         Affects_Academic_Performance)

write_csv(processed_data, "data/processed_data.csv")

head(processed_data)
```

The @sec-cleaning shows details of processing data:

-   Addicted_Score variable was replaced by Addicted_Level containing three level of social media addiction:

    -   1 to 3: Low

    -   4 to 6: Medium

    -   7 to 10: High

-   Affects_Academic_Performance variable was converted from “Yes/No” into "1/0" for statistical analysis purpose.

-   Incomplete records were removed from Processed_data.

### 4. Descriptive Statistics

The @fig-summary shows the number of students in each group of addiction level. The majority of examined students had moderate to high levels of social media addiction, with over 400 students and nearly 300 students, respectively. While the figure for the low group was only around 20.

```{r}
#| label: fig-summary
#| fig-cap: "Number of Students by Addiction Level"

library(ggplot2)

ggplot(processed_data, aes(x = Addicted_Level)) +
  geom_bar() +
  labs(
    x = "Addiction Level",
    y = "Count of Students"
  ) +
  theme_minimal()
```

The @tbl-summary summarise the mean, median, standard deviation, minimum, and maximum values for variables representing self-control of the entire sample.

-   Avg_Daily_Usage_Hours: Examined students have spent around 4.8 hours per day using social platforms. Minimum usage hour recorded is 1.5 hours, while the maximum is 8.5 hours, which indicates a clear difference.

-   Conflicts_Over_Social_Media: Students have a median of 3 conflicts with other people who are close to them. Although some do not have any conflict occurring between them and their loved one, some still have had 5 conflicts.

-   Affects_Academic_Performance: 64% of students of the entire sample agree that their studies have been impacted negatively by using social media.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of thre dependent variables"
library(tidyr)
library(knitr)
summary_vars <- processed_data %>%
  select(Avg_Daily_Usage_Hours, Conflicts_Over_Social_Media, Affects_Academic_Performance)

summary_table <- summary_vars %>%
  summarise(across(
    everything(),
    list(
      Mean = ~mean(.x, na.rm = TRUE),
      Median = ~median(.x, na.rm = TRUE),
      SD = ~sd(.x, na.rm = TRUE),
      Min = ~min(.x, na.rm = TRUE),
      Max = ~max(.x, na.rm = TRUE)
    )
  )) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("Variable", "Statistic"),
    names_pattern = "^(.*)_(Mean|Median|SD|Min|Max)$"
  ) %>%
  pivot_wider(
    names_from = Statistic,
    values_from = value
  )
kable(summary_table, digits = 2)
```

## 
